---
title: "gt-table"
author: "Jarred Robidoux"
date: "2/14/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# __Introduction__
This walkthrough will decipher how to utilize the **gt** library in R to make
tables for data visualization. 

# __Load required libraries__
```{r}
# install.packages("tidyverse")
library(tidyverse)
# install.packages("dplyr")
library(dplyr)
# install.packages("rvest")
library(rvest)
# install.packages("gt")
library(gt)
# install.packages("janitor")
library(janitor)
```
# __Scrape data from BasketballReference.com__
Data for this table will be provided from www.basketballreference.com
```{r}
# Upload the link
link <- "https://www.basketball-reference.com/leagues/NBA_2023_per_game.html"

full_table <- read_html(link, as.data.frame = T, stringsAsFactors = T)

# html_nodes to pull all nodes under the table label 
per_game <- full_table %>%
  html_nodes("table") %>%
  .[[1]] %>%
  html_table(fill = T)
```

# __Manipulate Data__
Before we create the table, we need to make changes to our dataset so that our table
can be easily readable. One of the first changes we have to make is to turn almost 
all of our values into numeric class instead of character. If this is not done, the 
data cannot be properly manipulated 
```{r}
per_game$PTS <- as.numeric(per_game$PTS)
per_game$AST <- as.numeric(per_game$AST)
per_game$`FG%` <- as.numeric(per_game$`FG%`)
per_game$FGA <- as.numeric(per_game$FGA)
per_game$`3P%` <- as.numeric(per_game$`3P%`)
per_game$`3PA` <- as.numeric(per_game$`3PA`)

per_game$`FG%` <-  per_game$`FG%` * 100
per_game$`3P%` <-  per_game$`3P%` * 100
```
Now we can use tidyverse to manipulate our data frame in order to make the 
table easier. 
```{r}
table_data <- per_game %>%
  select(Player, Pos, Tm, `FG%`, FGA, `3P%`, `3PA`, PTS) %>%
  arrange(desc(PTS)) %>%
  head(7)
```

# __Create Table__
```{r}
table_data %>%
  gt(rowname_col = "Player") %>%
  tab_header(
    title = md("**The League's Best Scorers**"),
    subtitle = md("`Players averaging 30+ PPG this season`")
  ) %>%
  fmt_number(columns = c("FGA", "FG%", "3P%", "3PA", "PTS"), decimals = 1)%>%
  cols_width(
    Pos ~ px(50),
    Tm ~ px(50),
    `FG%` ~ px(75),
    FGA ~ px(75),
    `3P%` ~ px(75),
    `3PA` ~ px(75),
    PTS ~ px(75)
  ) %>%
  tab_source_note(
    source_note = md("**Data : BasketballReference.com**")
  ) %>%
  tab_footnote(
    footnote = "Smallest Player averaging 30+ PPG (6'5)",
    locations = cells_stub(rows = "Shai Gilgeous-Alexander")) %>%
      opt_footnote_marks(marks = c("*", "**", "***"))
```

